---
title: "mSigHdp"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mSigHdp}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mSigHdp)
library(PCAWG7)
```

mSigHdp focuses on using hierarchical Dirichlet process mixture modeling
to discover mutational signatures. This is usually a very compute intensive, 
so here we will show only a toy example with tumors and the typical
classification of single-base-substitution (SBS) mutations in 96 types <ref>.
We take the mutational spectra from
10 tumors from the ICGC/TCGA Pan Cancer Analysis of Whole Genomes
Consortium (PCAWG).

```{r get_spectra}
toy_data <- PCAWG7::spectra$PCAWG$SBS96[ , 1:10]
colnames(toy_data) <- paste("T", 1:10, sep="")
toy_data[1:5,  ]
toy_data[91:96, ]
```

We set the _num.child.process_ and _CPU.cores_ to 4 for a quick run. 
*Add checkpoint*

```{r RunHdpxParallel}


retval <- mSigHdp::RunHdpxParallel(
  input.catalog       = toy_data,
  out.dir             = "output",
  num.child.process        = 4, 
  CPU.cores                = 4,
  seedNumber               = 123,
  K.guess                  = 5,
  burnin                   = 100,
  burnin.multiplier        = 2,
  post.n                   = 20, 
  post.space               = 10, 
  multi.types              = FALSE, 
  overwrite                = TRUE,
  gamma.alpha              = 1,
  gamma.beta               = 20, 
  high.confidence.prop     = 0.9,
  checkpoint               = TRUE,
  verbose                  = FALSE) 

dir()
dir("output")


```

