---
title: "Discovering mutational signatures with mSigHdp"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Discovering mutational signatures with mSigHdp}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mSigHdp)
library(ICAMS)
library(cosmicsig)
```

mSigHdp uses hierarchical Dirichlet process mixture modeling
to discover mutational signatures. 
For real data this is very compute intensive, 
so here we will show only a toy example with tumors and the typical
classification of single-base-substitution (SBS) mutations in 96 types. 
We generate the synthetic spectra based on two mutational
signatures, SBS1 and SBS22. We get the signatures from CRAN package,
which provides signatures from https://cancer.sanger.ac.uk/signatures/.
We set a specific seed so that the synthetic spectra are the 
same every time we run the vignette.


```{r get_spectra}
sigs <- cosmicsig::COSMIC_v3.2$signature$GRCh37$SBS96[  , c("SBS1", "SBS22")]
set.seed(2022)
```

We generate 10 random exposures and 2 highly-enriched exposures for
each signature (so we can discover them quickly),
and make them into an exposure matrix.

```{r get_spectra2}
n.tumor <- 10
exposures1 <- c(rnbinom(n.tumor, mu = 2000, size = 5), 10000, 100)
exposures2 <- c(rnbinom(n.tumor, mu = 10000, size = 1000), 100, 5000)
exposures <- rbind(exposures1, exposures2)
```

Then we generate spectra by multiplying the signatures times the
exposures and rounding, because we need integer mutation counts.
We convert the matrix into an ICAMS catalog so we can use custom plotting later.

```{r get_spectra3}
toy_data <- round(sigs %*% exposures)
colnames(toy_data) <- paste("T", 1:(n.tumor + 2), sep="")
toy_data <- ICAMS::as.catalog(toy_data, catalog.type = "counts")
```

Here are the first few and last few rows of the 12 spectra.
Each column is one of the 12 synthetic tumors.
Each row is a mutation type, for example in in row 5
the row name CCAA indicates a mutation from CCA to CAA.

```{r show_spectra}
knitr::kable(toy_data[1:5,  ])
knitr::kable(toy_data[91:96, ])
```

Must people will want to use the `RunHdxParallel` function.
*Most of the parameters in the example here are not suitable for real data*.
Please see our paper to be posted on bioRxiv soon for suggestions
for actual values to use.

```{r RunHdpxParallel}

retval <- mSigHdp::RunHdpxParallel(
  input.catalog        = toy_data,
  out.dir              = "vignette_output",
  num.child.process    = 4, 
  CPU.cores            = 4,
  seedNumber           = 123,
  K.guess              = 5,
  burnin               = 100,
  burnin.multiplier    = 2,
  post.n               = 5, 
  post.space           = 10, 
  multi.types          = FALSE, 
  overwrite            = TRUE,
  gamma.alpha          = 1,
  gamma.beta           = 20, 
  high.confidence.prop = 0.9,
  checkpoint           = TRUE,
  verbose              = FALSE) 
```

The results are in both `retval` and in subdirectory `vignette_output`,
which also contains plots of the output and diagnostic plots. 
It is convenient to automatically save the results to files
since the program will likely have been running for days, and you will
likely want to save the outputs to files in any case.
The return value of `RunHdpParallel` will be included in the
directory in binary form as `hdp.retval.Rdata`.

The directory contains a plot of the extracted (discovered) signatures
in file `extracted.signatures.pdf`.

```{r sigplot, out.width="100%", out.height="50%", echo=F}
knitr::include_graphics("vignette_output/extracted.signatures.pdf")
```

The signatures in numerical form are in file `extracted.signatures.csv`.

```{r extracted_sigs}
# dir()
# dir("vignette_output")
extracted.sigs <- ICAMS::ReadCatalog("vignette_output/extracted.signatures.csv")
knitr::kable(extracted.sigs[1:10, ])

```




